---
title             : "Correlation of trace metal and Depression"
shorttitle        : "Trace metal and Depression"

author: 
  - name          : "Hao Chen"
    corresponding : yes    # Define only one corresponding author
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"


affiliation:
  - id            : "1"
    institution   : "University of Chicago"


abstract: |
  This study explores the link between Major Depressive Disorder (MDD) and essential trace elements, analyzing data from the National Health and Nutrition Examination Survey (NHANES). MDD, affecting an estimated 280 million people worldwide, manifests through diverse symptoms impacting emotional, physical, and cognitive well-being. The focus is on trace elements necessary for metabolic functions, as alterations in these elements are implicated in MDD's development. Utilizing NHANES data, the study examines urine samples from participants aged 3 and above for trace metal analysis, employing mass spectrometry for precise quantification. Depressive symptoms are assessed via the PHQ-9 questionnaire, a recognized depression screening tool. By investigating the relationship between trace metals and MDD symptoms, this research aims to enhance our understanding of MDD's etiology and inform potential treatment approaches.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "Trace metal, Depression"
wordcount         : "X"

bibliography      : ["./nhanes2017.bib"]

floatsintext      : yes
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
header-includes:
  - \usepackage{colortbl}
---

```{r setup, include = FALSE}
library("papaja")
library(readr)
library(dplyr)
library(ggplot2)
library(knitr)
library(kableExtra)
r_refs("./nhanes2017.bib")
nhanes_data <- read_csv("~/Documents/Github/Env_chem-cortisol/nhanes2017.csv")
source("~/Documents/GitHub/Env_chem-cortisol/data-prep.R")
knitr::opts_chunk$set(warning=FALSE)
```

# Introduction
Major Depressive Disorder (MDD) is a prevalent yet severe mood condition marked by experience of low mood and negative emotions for a long period of time[@americanpsychiatricassociationDiagnosticStatisticalManual2013].  In 2019, approximately 280 million individuals, which includes 5% of the adult population, were estimated to have experienced depression[@GBDResults]. According to the statistics from the United States National Institute of Mental Health, around 21.0 million adults in the United States experienced at least one major depressive episode, accounting for 8.3% of all U.S. adults[@MajorDepressionNational].
MDD is a multifaceted and intricate condition, which can result in impairment of psychosocial functioning and quality of life[@saragoussiLongtermFollowupHealthrelated2018]. In addition to depressed feelings, patients with MDD may experience a wide range of physical and cognitive symptoms, including feelings of sadness, irritability, loss of interest or pleasure in activities, changes in appetite or weight, sleep disturbances, fatigue, feelings of worthlessness or guilt, difficulty concentrating, and thoughts of death or suicide[@americanpsychiatricassociationDiagnosticStatisticalManual2013]. 
Etiology of MDD includes biological, environmental, and personal vulnerabilities[@nationalresearchcouncilusandinstituteofmedicineuscommitteeondepressionEtiologyDepression2009]. Lately, there has been growing attention towards metallomic research in psychiatry, with a focus on examining the involvement of essential trace elements in both the development and progression of MDD symptoms. An essential trace element refers to a mineral or dietary element necessary in small amounts for an organism’s proper growth, development, and physiology[@bowen1966trace]. These elements are vital for conducting essential metabolic activities in organisms. Examples of essential trace metals in human nutrition include Na, K, Mg, Ca, Fe, Mn, Co, Cu, Zn and Mo[@zorodduEssentialMetalsHumans2019]. The trace metals play important catalytic and structural roles. These elements facilitate essential biochemical reactions by serving as cofactors for numerous enzymes and as stabilizing agents for the structures of enzymes and proteins[@prashanthReviewRoleEssential2015]. Alterations in the accumulation or absence of these components can trigger alternative metabolic pathways, potentially contributing to various neurodevelopmental diseases and conditions [@yuiEditorialThematicIssue2016].
Baj et al. (2013) conducted an narrative review of the relationship between levels of selected trace elements in the serum of individuals with MDD and the initiation and advancement of this mental health disorder[@bajTraceElementsLevels2023]. Findings of this review reveal that the levels of metal content in the body are related to the outcomes of individuals with MDD in various ways. For example, Li et al. (2020) has demonstrated that elevated levels of copper can disrupt the functioning of NMDA receptors, contributing to cognitive deficits in MDD[@liAlleviationCognitiveDeficits2020]. Increased copper concentrations can also impair AMPA receptor function, leading to disruptions in glutamatergic transmission, supporting the Glu hypothesis of depression[@gerhardEmergingTreatmentMechanisms2016; @petersBiphasicEffectsCopper2011; @styczenSerumZincConcentration2017].
Although the relationship between the trace metals and MDD has been widely studied, there are limited research on the results of trace metal and how they relate to the onset of MDD symptom. This study aims to reveal the relationship between the two.


# Methods

## Dataset

This research employs a cross-sectional methodology, leveraging data from the National Health and Nutrition Examination Survey (NHANES) conducted by the US National Center for Health Statistics. Objectives of the survey encompass evaluating the health and nutritional conditions of individuals across the United States, as well as identifying the prevalence of significant diseases and their risk contributors. The NHANES database includes a wide array of information, such as demographic specifics, nutritional insights, results from physical exams, laboratory test results, participant questionnaire answers, and confidential data.Here is a table showing the racial demographic of our study.
```{r racial-demographic}
library(knitr)
library(kableExtra)

# Create a lookup table for racial categories
racial_lookup <- data.frame(
  RIDRETH3 = c(1, 2, 3, 4, 6, 7, NA),
  Racial_Category = c("Mexican American", "Other Hispanic", "Non-Hispanic White", 
                      "Non-Hispanic Black", "Non-Hispanic Asian", "Other", "Missing")
)

# Use the lookup table to create a new column with racial categories
nhanes_data$Racial_Category <- racial_lookup$Racial_Category[match(nhanes_data$RIDRETH3, racial_lookup$RIDRETH3)]

# Use the lookup table to create a new column with racial categories
nhanes_data$Racial_Category <- racial_lookup$Racial_Category[match(nhanes_data$RIDRETH3, racial_lookup$RIDRETH3)]

# Create a table showing the racial demographics
racial_demographics <- table(nhanes_data$Racial_Category)

# Print the formatted table with kable
kable(racial_demographics, format = "latex", caption = "Racial Demographics") %>%
  kable_styling(latex_options = c("striped", "scale_down"))


```
## Participants

Participants aged 3 to 5 years, along with a one-third subset of those aged 6 and above, were considered eligible for this study. Due to privacy concerns, access to urine lead data for the 3 to 5 age group and urine strontium and uranium data for those aged 3 and above is restricted to the NCHS Research Data Center. However, the dataset does include urine lead data for participants aged 6 and older, as well as urine barium, cadmium, cesium, cobalt, manganese, molybdenum, antimony, thallium, tin, and tungsten for all eligible participants aged 3 and above. For further details on accessing urine lead, strontium, and uranium data, refer to the Analytic Notes.

## Description of Laboratory Methodology for Trace Metal

This technique accurately quantifies various metals in urine samples by utilizing mass spectrometry, preceded by a straightforward dilution preparation of the samples. The process begins with the introduction of liquid specimens into the mass spectrometer via an inductively coupled plasma (ICP) ionization source. Here, a nebulizer converts the sample into fine droplets within an argon gas stream. These droplets then proceed into the ICP, where they are ionized. The ions navigate through a focusing area, enter the dynamic reaction cell (DRC), pass through the quadrupole mass filter, and ultimately, the detector sequentially counts them in rapid succession. This method enables the precise identification of individual isotopes for each element analyzed.

## Detection Limits for Trace Metal

The detection limits  remained uniform across all analytes within the dataset. For each analyte, two specific variables are furnished. A variable name ending with “LC” (for example, URDUBALC) signifies if the measurement fell below the detection limits : a "0" value indicates the measurement was at or above this limits, whereas a "1" denotes it was below. Conversely, the variable starting with URX (for example, URXUBA) reports the actual measurement for the analyte. When the measurement for an analyte is below the detection limit (for instance, URDUBALC=1), a predetermined substitute value is used in its place. This substitute value is calculated as the detection limit divided by the square root of 2.

## Depression Assessment 

The assessment of participants' depressive symptoms was conducted using the Patient Health Questionnaire-9 (PHQ-9), a nine-item instrument designed for depression screening. The PHQ-9 is recognized as a reliable and validated method for detecting common depression and related disorders, especially in primary care environments[@kimUrinaryPhthalateMetabolites2016]. This questionnaire includes nine prompts, with responses scored based on the Diagnostic and Statistical Manual of Mental Disorders (DSM) criteria as follows: 0 (“Never”), 1 (“A few days”), 2 (“More than a week”), and 3 (“Almost every day”), leading to a total possible score between 0 and 27. A score within the 0-9 range is considered indicative of a non-depressive state, whereas a score of 10 or higher suggests the presence of depression. The PHQ-9 has demonstrated high sensitivity (0.88) and specificity (0.85) for identifying depression at a threshold score of 10 or above[@levisAccuracyPatientHealth2019].

## Data analysis
We used `r cite_r("./nhanes2017.bib")` for all our analyses. With in R, we performed Spearman's Rank Correlation coefficient test to see the relationship between the sum of over-limit trace metal type(LCsum) and their PHQ-9 score(DPQsum) since both our variables are ordinal variables. Besides that, we also performed Spearman's Rank Correlation coefficient test for each individual trace metal. 

# Results
```{r descriptive analysis, echo=FALSE,results='hide'}
#For my descriptive analysis. I will show the distribution of LCsum and DPQsum, the means, median.
# LCDPQ_data: Exclude NA in LCsum, ensure LCsum is not 0, and DPQsum is above 10
LCDPQ_data <- fix_nhanes_data[!is.na(fix_nhanes_data$LCsum) & 
                                 fix_nhanes_data$LCsum != 0 & 
                                 fix_nhanes_data$DPQsum > 10, ]
# Summary statistics for LCsum
summary(LCDPQ_data$LCsum)

# Calculate mean
mean_LCsum <- mean(LCDPQ_data$LCsum, na.rm = TRUE)

# Calculate median
median_LCsum <- median(LCDPQ_data$LCsum, na.rm = TRUE)


# Summary statistics for DPQsum
summary(LCDPQ_data$DPQsum)

# Calculate mean for DPQsum
mean_DPQsum <- mean(LCDPQ_data$DPQsum, na.rm = TRUE)

# Calculate median for DPQsum
median_DPQsum <- median(LCDPQ_data$DPQsum, na.rm = TRUE)


```
In our descriptive analysis, we examined the distribution, mean, and median values of two key variables: the sum of trace metal concentrations exceeding detection limits (LCsum) and the PHQ-9 score (DPQsum), which quantifies the frequency of depression symptoms. The summary statistics for LCsum revealed a mean value of `r mean_LCsum` and a median value of `r median_LCsum`, indicating the central tendency of trace metal concentrations in our study population. Similarly, for DPQsum, the mean score was calculated to be `r mean_DPQsum`, with a median score of `r median_DPQsum`, providing insight into the prevalence and distribution of depression symptoms among participants.


This is a plot showing the correlation of LCsum and DPQsum. As shown in the plot \@ref(fig:LCDPQ-plot), we see observe a relationship between the sum of trace metal concentrations exceeding detection limits (LCsum) and the frequency of depression symptoms as quantified by the PHQ-9 score (DPQsum) for individuals with DPQsum scores greater than 10. The scatter plot, enhanced by jitter to reduce overplotting, and the linear regression line suggest a slightly positive trend, indicating that as the LCsum increases, the DPQsum slight increases. This visual analysis is supported by a slight slope of the regression line, colored in blue. Although the plot reveals the general trend, it is essential to consider the variability around the regression line. 

```{r LCDPQ-plot, fig.cap="LCsum vs DPQsum for Individuals with DPQsum > 10"}
library(ggplot2)

LCDPQ_plot <- ggplot(LCDPQ_data, aes(x = LCsum, y = DPQsum)) +
  geom_jitter() +  
  geom_smooth(method = "lm", color = "blue") +  
  labs(title = "LCsum vs DPQsum for Individuals with DPQsum > 10",
       x = "LCsum", y = "DPQsum")

print(LCDPQ_plot)

```


```{r hypothesis test analysis}
#For my hypothesis test analysis. I would do a Spearman's Rank Correlation coefficient since both my variables are ordinal variables.One is LCsum. It is the sum of total kinds of trace metal above detection limits in each participants. One is DPQsum, which is their depression test result.

spearman_test_LCsum <- cor.test(LCDPQ_data$LCsum, LCDPQ_data$DPQsum, method = "spearman")
# r=0.18 p=0.04
```
The Spearman test result for the type of trace metal exceed detection limits and the frequency of depression symptom score shows a correlation coefficient (rho) of `r round(spearman_test_LCsum$estimate, 4)` with a p-value of `r round(spearman_test_LCsum$p.value, 4)`.The Spearman's rho value of 0.18 indicates a weak positive correlation between the sum of trace metals detected in participants and their scores on the depression test. This suggests that there is a slight tendency for individuals with a higher number of trace metal types above detection limits to report higher levels of depressive symptoms. However, the correlation strength is weak, implying that while there is a positive relationship, it is not particularly strong.The p-value of 0.04 is less than the commonly used significance level of 0.05. This indicates that the observed correlation is statistically significant.

```{r test-result-table}

# Specify the variables for which to perform Spearman's rank correlation test
variables <- c("URXUBA", "URXUCD", "URXUCO", "URXUCS", "URXUMO",
               "URXUMN", "URXUPB", "URXUSB", "URXUSN", "URXUTL", "URXUTU")

# Initialize an empty data frame to store results
results <- data.frame(Variable = character(),
                      R_Score = numeric(),
                      P_Value = numeric(),
                      stringsAsFactors = FALSE)

# Loop through each variable and perform the test
for (var in variables) {
  test_result <- cor.test(LCDPQ_data[[var]], LCDPQ_data$DPQsum, method = "spearman")
  
# Append the results to the results dataframe
  results <- rbind(results, data.frame(Variable = var,
                                       R_Score = test_result$estimate,
                                       P_Value = test_result$p.value))
}
knitr::kable(results, caption = "Individual Trace Metal Spearman's Rank Correlation Test Results")


```

The result for each type of trace metal is stored in table 2. The results indicate that most trace metals have a weak to moderate association with depression symptoms, with many not reaching statistical significance. Notably, URXUMN and URXUSB show statistically significant positive correlations, suggesting that higher concentrations of these metals are associated with more frequent depression symptoms.

Manganese (URXUMN, rho: 0.2216, p-value: 0.0116): The analysis revealed a moderate positive correlation between manganese levels and the frequency of depression symptoms that is statistically significant. This suggests that higher manganese exposures are associated with an increased frequency of reported depression symptoms among the study participants. Given the statistical significance (p < 0.05), this finding indicates a meaningful association between manganese exposure and mental health outcomes. It underscores the need for further investigation into the biological mechanisms through which manganese may influence depressive symptoms and to explore potential public health interventions to mitigate its negative impacts on mental health.

Antimony (URXUSB, rho: 0.1815, p-value: 0.0395): The results indicate a moderate positive correlation between antimony levels and the frequency of depression symptoms that is also statistically significant. Similar to manganese, this finding suggests an association between higher antimony exposures and an increased frequency of depression symptoms. The statistical significance of this correlation points to antimony as another trace metal of concern in the context of mental health, warranting additional research to understand its effects on depression and to assess the risk it may pose to exposed populations.

# Discussion
This study explored the association between trace metal exposure, quantified as the sum of trace metals exceeding detection limits (LCsum), and the severity of depressive symptoms, measured using the PHQ-9 score (DPQsum), in a sample population. Our findings reveal a weak but statistically significant positive correlation between LCsum and DPQsum, suggesting that increased exposure to trace metals may be associated with higher levels of depressive symptoms. Specifically, the analysis identified manganese (URXUMN) and antimony (URXUSB) as having moderate and statistically significant positive correlations with depression symptoms, highlighting their potential roles as environmental risk factors for mental health.The observed association between trace metal exposure and depression aligns with emerging literature that suggests environmental pollutants, including heavy metals, may contribute to the pathophysiology of depressive disorders through various biological mechanisms. 

## Limitation
This study's limitations include its cross-sectional design, which precludes causal inferences. While significant associations were observed, it is not possible to determine whether metal exposure precedes the development of depressive symptoms or vice versa. Additionally, reliance on self-reported depression scores may introduce reporting biases. The measurement of metal exposure was based on the sum of metals exceeding detection limits, which may not fully capture the nuances of individual metal exposures or their combined effects.

## Future Directions
Longitudinal studies are needed to establish temporality and causality in the relationship between trace metal exposure and depression. Further research should also explore the biological mechanisms underlying the observed associations, with particular attention to low-level exposures and potential synergistic effects of multiple metals. Public health interventions aimed at reducing exposure to identified metals, such as manganese and antimony, could be evaluated for their impact on mental health outcomes.

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
