---
title             : "Correlation of trace metal and frequency of depression symptom"
shorttitle        : "Title"

author: 
  - name          : "HaoChen"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "University of Chicago"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.1

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  One or two sentences to put the results into a more **general context**.
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "Trace metal, Depression"
wordcount         : "X"

bibliography      : ["./nhanes2017.bib"]

floatsintext      : no
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---
#Introduction
Major Depressive Disorder (MDD) is a prevalent yet severe mood condition marked by experience of low mood and negative emotions for a long period of time[@americanpsychiatricassociationDiagnosticStatisticalManual2013].  In 2019, approximately 280 million individuals, which includes 5% of the adult population, were estimated to have experienced depression[@GBDResults]. According to the statistics from the United States National Institute of Mental Health, around 21.0 million adults in the United States experienced at least one major depressive episode, accounting for 8.3% of all U.S. adults[@MajorDepressionNational].
MDD is a multifaceted and intricate condition, which can result in impairment of psychosocial functioning and quality of life[@saragoussiLongtermFollowupHealthrelated2018]. In addition to depressed feelings, patients with MDD may experience a wide range of physical and cognitive symptoms, including feelings of sadness, irritability, loss of interest or pleasure in activities, changes in appetite or weight, sleep disturbances, fatigue, feelings of worthlessness or guilt, difficulty concentrating, and thoughts of death or suicide[@americanpsychiatricassociationDiagnosticStatisticalManual2013]. 
Etiology of MDD includes biological, environmental, and personal vulnerabilities[@nationalresearchcouncilusandinstituteofmedicineuscommitteeondepressionEtiologyDepression2009]. Lately, there has been growing attention towards metallomic research in psychiatry, with a focus on examining the involvement of essential trace elements in both the development and progression of MDD symptoms. An essential trace element refers to a mineral or dietary element necessary in small amounts for an organism’s proper growth, development, and physiology[@bowen1966trace]. These elements are vital for conducting essential metabolic activities in organisms. Examples of essential trace metals in human nutrition include Na, K, Mg, Ca, Fe, Mn, Co, Cu, Zn and Mo[@zorodduEssentialMetalsHumans2019]. The trace metals play important catalytic and structural roles. These elements facilitate essential biochemical reactions by serving as cofactors for numerous enzymes and as stabilizing agents for the structures of enzymes and proteins[@prashanthReviewRoleEssential2015]. Alterations in the accumulation or absence of these components can trigger alternative metabolic pathways, potentially contributing to various neurodevelopmental diseases and conditions [@yuiEditorialThematicIssue2016].
Baj et al. (2013) conducted an narrative review of the relationship between levels of selected trace elements in the serum of individuals with MDD and the initiation and advancement of this mental health disorder[@bajTraceElementsLevels2023]. Findings of this review reveal that the levels of metal content in the body are related to the outcomes of individuals with MDD in various ways. For example, Li et al. (2020) has demonstrated that elevated levels of copper can disrupt the functioning of NMDA receptors, contributing to cognitive deficits in MDD[@liAlleviationCognitiveDeficits2020]. Increased copper concentrations can also impair AMPA receptor function, leading to disruptions in glutamatergic transmission, supporting the Glu hypothesis of depression[@gerhardEmergingTreatmentMechanisms2016; @petersBiphasicEffectsCopper2011; @styczenSerumZincConcentration2017].
Although there relationship between the trace metals in human serum and MDD has been widely studied, there are limited research on the results of trace metal and how they relate to the frequency of MDD symptom. This study aims to reveal the relationship between the two.

```{r setup, include = FALSE}
library("papaja")
library(readr)
library(dplyr)
r_refs("./nhanes2017.bib")
nhanes_data <- read_csv("~/Documents/Github/Env_chem-cortisol/nhanes2017.csv")
source("~/Documents/GitHub/Env_chem-cortisol/data-prep.R")
```

Here to reference my depression measure gender difference plot.(Figure\@ref(fig:gender_diff_plot))
```{r gender_diff_plot}
#Create a bar plot to visualize the difference in mean depression scores (DPQsum) between genders
#box plot would be better. add standard error bar if want to leave as bar plot.
library(dplyr)
library(ggplot2)
fix_nhanes_data <- fix_nhanes_data %>%
  mutate(Gender = if_else(RIAGENDR == 1, "Male", "Female"))
gender_diff_plot<-ggplot(fix_nhanes_data, aes(x = Gender, y = DPQsum)) +
  geom_boxplot(position = "dodge") +
  labs(title = "Difference in Mean DPQsum Between Genders",
       x = "Gender",
       y = "Mean DPQsum")
print(gender_diff_plot)
# To "fine tune" it, I can change the color of two bar and make it a little thinner. 1
```
Here to reference my LCsum and DPQ sum linear regression plot. (Figure\@ref(fig:LCDPQ_plot))
```{r LCDPQ-plot}
#My second plot would be a linear regression 
library(ggplot2)
LCsumnotzero_data <- fix_nhanes_data[!is.na(fix_nhanes_data$LCsum) & fix_nhanes_data$LCsum != 0, ]
LCDPQ_plot <- ggplot(LCsumnotzero_data, aes(x = LCsum, y = DPQsum)) +
  geom_jitter() +
   geom_smooth(method = "lm", color = "blue") +
  labs(title = "LCsum vs DPQsum", x = "LCsum", y = "DPQsum")
print(LCDPQ_plot)
```


Here to reference my racial demographic table.(Table:\@ref(racial-demographic))
```{r racial-demographic}
library(knitr)
library(kableExtra)

# Create a lookup table for racial categories
racial_lookup <- data.frame(
  RIDRETH3 = c(1, 2, 3, 4, 6, 7, NA),
  Racial_Category = c("Mexican American", "Other Hispanic", "Non-Hispanic White", 
                      "Non-Hispanic Black", "Non-Hispanic Asian", "Other", "Missing")
)

# Use the lookup table to create a new column with racial categories
nhanes_data$Racial_Category <- racial_lookup$Racial_Category[match(nhanes_data$RIDRETH3, racial_lookup$RIDRETH3)]

# Use the lookup table to create a new column with racial categories
nhanes_data$Racial_Category <- racial_lookup$Racial_Category[match(nhanes_data$RIDRETH3, racial_lookup$RIDRETH3)]

# Create a table showing the racial demographics
racial_demographics <- table(nhanes_data$Racial_Category)

# Print the formatted table with kable
kable(racial_demographics, format = "latex", caption = "Racial Demographics") %>%
  kable_styling(latex_options = c("striped", "scale_down"))


```
# Methods

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("./nhanes2017.bib")` for all our analyses.
```{r hypothesis test analysis}
#For my hypothesis test analysis. I would do a Spearman's Rank Correlation coefficient since both my variables are ordinal variables.One is LCsum. It is the sum of total kinds of trace metal above detection limits in each participants. One is DPQsum, which is their depression test result.

spearman_test_LCsum <- cor.test(fix_nhanes_data$LCsum, fix_nhanes_data$DPQsum, method = "spearman")
print(spearman_test_LCsum)#r=0.0303 p=0.191
```
The spearman test result for the type of trace metal exceed detection limits and the frequency of depression symptom is `r print(spearman_test_LCsum)`
```{r table for test result}

# Specify the variables for which to perform Spearman's rank correlation test
variables <- c("URXUBA", "URXUCD", "URXUCO", "URXUCS", "URXUMO",
               "URXUMN", "URXUPB", "URXUSB", "URXUSN", "URXUTL", "URXUTU")

# Initialize an empty data frame to store results
results <- data.frame(Variable = character(),
                      R_Score = numeric(),
                      P_Value = numeric(),
                      stringsAsFactors = FALSE)

# Loop through each variable and perform the test
for (var in variables) {
  test_result <- cor.test(fix_nhanes_data[[var]], fix_nhanes_data$DPQsum, method = "spearman")
  
# Append the results to the results dataframe
  results <- rbind(results, data.frame(Variable = var,
                                       R_Score = test_result$estimate,
                                       P_Value = test_result$p.value))
}

print(results)

```
```{r descriptive analysis}
#For my descriptive analysis. I will show the distribution of LCsum and DPQsum, the means, median.
# Summary statistics for LCsum
summary(fix_nhanes_data$LCsum)

# Calculate mean
mean_LCsum <- mean(fix_nhanes_data$LCsum, na.rm = TRUE)
print(paste("Mean of LCsum:", mean_LCsum))

# Calculate median
median_LCsum <- median(fix_nhanes_data$LCsum, na.rm = TRUE)
print(paste("Median of LCsum:", median_LCsum))

# Summary statistics for DPQsum
summary(fix_nhanes_data$DPQsum)

# Calculate mean for DPQsum
mean_DPQsum <- mean(fix_nhanes_data$DPQsum, na.rm = TRUE)
print(paste("Mean of DPQsum:", mean_DPQsum))

# Calculate median for DPQsum
median_DPQsum <- median(fix_nhanes_data$DPQsum, na.rm = TRUE)
print(paste("Median of DPQsum:", median_DPQsum))

```
Means of LCsum is `r print(paste("Median of DPQsum:", median_DPQsum))`, which means participants have an average of 9 or 10 types of trace metal above detection limits. For the same people group, their average score for frequency of depression symptom is `r print(paste("Mean of DPQsum:", mean_DPQsum))`. 
# Results

# Discussion


\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
