---
title             : "Urine_Dep"
shorttitle        : "Title"

author: 
  - name          : "HaoChen"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.1

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  One or two sentences to put the results into a more **general context**.
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["nhanes2017.bib"]

floatsintext      : no
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("nhanes2017.bib")
source("~/Documents/GitHub/Env_chem-cortisol/data-prep.R")
```


```{r}
#selected the depression related data and write it into a individual csv file
library(readr)
library(dplyr)
nhanes_data <- read_csv("~/Documents/Github/Env_chem-cortisol/nhanes2017.csv")
depression_columns <- nhanes_data %>% select(starts_with("DPQ"),starts_with("SEQN"))
head(depression_columns)
write_csv(depression_columns,"depression_columns")

```
Here to reference my depression measure gender difference plot.Figure\\@ref(fig:DPQgenderdiff-plot)
```{r DPQgenderdiff-plot}
#Create a bar plot to visualize the difference in mean depression scores (DPQsum) between genders
#box plot would be better. add standard error bar if want to leave as bar plot.
library(dplyr)
library(ggplot2)
fix_nhanes_data <- fix_nhanes_data %>%
  mutate(Gender = if_else(RIAGENDR == 1, "Male", "Female"))
gender_diff_plot<-ggplot(fix_nhanes_data, aes(x = Gender, y = DPQsum)) +
  geom_boxplot(position = "dodge") +
  labs(title = "Difference in Mean DPQsum Between Genders",
       x = "Gender",
       y = "Mean DPQsum")

# To "fine tune" it, I can change the color of two bar and make it a little thinner. 1
```
Here to reference my LCsum and DPQ sum linear regression plot. Figure\\@ref(fig:LCDPQ-plot)
```{r LCDPQ-plot}
#My second plot would be a linear regression 
library(ggplot2)
LCsumnotzero_data <- fix_nhanes_data[!is.na(fix_nhanes_data$LCsum) & fix_nhanes_data$LCsum != 0, ]
ggplot(LCsumnotzero_data, aes(x = LCsum, y = DPQsum)) +
  geom_jitter() +
   geom_smooth(method = "lm", color = "blue") +
  labs(title = "LCsum vs DPQsum", x = "LCsum", y = "DPQsum")

```


Here to reference my racial demographic table @ref(racial_demographics)
```{r racial_demographics}
library(knitr)
library(kableExtra)
# Create a lookup table for racial categories
racial_lookup <- data.frame(
  RIDRETH3 = c(1, 2, 3, 4, 6, 7, NA),
  Racial_Category = c("Mexican American", "Other Hispanic", "Non-Hispanic White", 
                      "Non-Hispanic Black", "Non-Hispanic Asian", "Other", "Missing")
)

# Use the lookup table to create a new column with racial categories
nhanes_data$Racial_Category <- racial_lookup$Racial_Category[match(nhanes_data$RIDRETH3, racial_lookup$RIDRETH3)]

# Use the lookup table to create a new column with racial categories
nhanes_data$Racial_Category <- racial_lookup$Racial_Category[match(nhanes_data$RIDRETH3, racial_lookup$RIDRETH3)]

# Create a table showing the racial demographics
racial_demographics <- table(nhanes_data$Racial_Category)

# Print the table with kable and style it with kableExtra
rac_table <- kable(racial_demographics, col.names = c("Race", "Count"), align = "c", caption = "Racial Demographics in NHANES Data") %>%
  kable_styling(latex_options = "striped", position = "center") %>%
  column_spec(1, bold = TRUE, border_right = TRUE) %>%
  row_spec(0, bold = TRUE)

# Print the formatted table
print(rac_table)


```
# Methods

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("nhanes2017.bib")` for all our analyses.
```{r hypothesis test analysis}
#For my hypothesis test analysis. I would do a Spearman's Rank Correlation coefficient since both my variables are ordinal variables.One is LCsum. It is the sum of total kinds of trace metal above detection limits in each participants. One is DPQsum, which is their depression test result.

spearman_test_LCsum <- cor.test(fix_nhanes_data$LCsum, fix_nhanes_data$DPQsum, method = "spearman")
print(spearman_test_LCsum)#r=0.0303 p=0.191
```

```{r table for test result}

# Specify the variables for which to perform Spearman's rank correlation test
variables <- c("URXUBA", "URXUCD", "URXUCO", "URXUCS", "URXUMO",
               "URXUMN", "URXUPB", "URXUSB", "URXUSN", "URXUTL", "URXUTU")

# Initialize an empty data frame to store results
results <- data.frame(Variable = character(),
                      R_Score = numeric(),
                      P_Value = numeric(),
                      stringsAsFactors = FALSE)

# Loop through each variable and perform the test
for (var in variables) {
  test_result <- cor.test(fix_nhanes_data[[var]], fix_nhanes_data$DPQsum, method = "spearman")
  
# Append the results to the results dataframe
  results <- rbind(results, data.frame(Variable = var,
                                       R_Score = test_result$estimate,
                                       P_Value = test_result$p.value))
}

print(results)

```
```{r descriptive analysis}
#For my descriptive analysis. I will show the distribution of LCsum, the means, median.
```
# Results

# Discussion


\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
