---
title             : "Urine_Dep"
shorttitle        : "Title"

author: 
  - name          : "HaoChen"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.1

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  One or two sentences to put the results into a more **general context**.
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
source("~/Documents/GitHub/Env_chem-cortisol/Assignment2-1.R")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)

```
```{r}

# Call the function
generate_message <- helloworld("Ricky", "Chen", "Male")

# Print the greeting message
print(generate_message)
```


```{r-sameple_urxumn}

#read data and get 10 sample data from column "URXUMN
library(readr)
library(dyplr)
nhanes_data <- read_csv("~/Documents/Github/Env_chem-cortisol/nhanes2017.csv")
sample_data_10 <- sample_n(nhanes_data,10,replace = TRUE)
sampled_urxumn <- sample(nhanes_data$URXUMN, 10, replace = TRUE)
print(sample_urxumn)
```
```{r}
#selected the depression related data and write it into a individual csv file
library(readr)
library(dplyr)
nhanes_data <- read_csv("~/Documents/Github/Env_chem-cortisol/nhanes2017.csv")
depression_columns <- nhanes_data %>% select(starts_with("DPQ"),starts_with("SEQN"))
head(depression_columns)
write_csv(depression_columns,"depression_columns")

```

```{r}
#Create a bar plot to visualize the difference in mean depression scores (DPQsum) between genders
#box plot would be better. add standard error bar if want to leave as bar plot.
library(dplyr)
library(ggplot2)
fix_nhanes_data <- fix_nhanes_data %>%
  mutate(Gender = if_else(RIAGENDR == 1, "Male", "Female"))
ggplot(fix_nhanes_data, aes(x = Gender, y = DPQsum)) +
  geom_boxplot(position = "dodge") +
  labs(title = "Difference in Mean DPQsum Between Genders",
       x = "Gender",
       y = "Mean DPQsum")
# To "fine tune" it, I can change the color of two bar and make it a little thinner. 1
```

```{r}
#My second plot would be a linear regression 
library(ggplot2)
ggplot(fix_nhanes_data, aes(x = LCsum, y = DPQsum)) +
  geom_jitter() +
   geom_smooth(method = "lm", color = "blue") +
  labs(title = "LCsum vs DPQsum", x = "LCsum", y = "DPQsum")

```
# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

<!-- This paragraph is for practice -->
**This line should be bold**g
*This line should be italicized*
[link](www.4399.com)
*Unorder list
*Unorder list 2
1.Order list
2.Order list 2
## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.


# Results

# Discussion


\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
